Query Optimization for Dynamic Imputation

### 2017-07-14 paper/algorithm.tex
minor change in table language
paper: add cameraready

\begin{table}
  \centering
  \caption{Notation used in Section~\ref{sec:algorithm}.}
  \begin{tabular}{ll}
    \toprule
    Operator & Description \\
    \midrule
    $\textsc{Attr}(\cdot)$ & Attributes used in a predicate. \\
    $\textsc{Time}(\cdot)$ & Estimated time to run a plan. \\
    $\textsc{Penalty}(\cdot)$ & Quality penalty of a query plan. \\
    $\textsc{Card}(\cdot)$ & Estimated cardinality of a plan. \\
    $\widehat{Var}(\cdot)$ & Estimated variance of a column. \\
    $\textsc{Dirty}(\cdot)/\textsc{Clean}(\cdot)$ & Dirty/clean attributes in a plan. \\
    $Q[\cdot]$ & Plan cache access operator. \\
    $\cdot \triangleleft \cdot$ & Plan cache insertion operator. \\
    \bottomrule
  \end{tabular}
\end{table}

@@

Separately, we specify a heuristic penalty that takes into account the
improvement in imputation quality from more tuples and attributes, as well as
the diminishing returns from more data.
\[
    P(q) = \frac{|\cameraready{C}|}{\sqrt{ \textsc{Attr}(q) \times \textsc{Card}(q)}}
\]

@@

We can take advantage of the histograms maintained by \ProjectName{} to estimate the mean and construct a non-blocking impute operator.
Applying the heuristics from~\Cref{sec:quality} and~\Cref{sec:query-runtime}, we
have
\begin{equation*}
\begin{aligned}
    P(q) &= \sum_{c \in \cameraready{C}} \frac{1+\widehat{Var}(c)}{\textsc{Card}(q)} \\
    T(q) &\approx 0.
\end{aligned}
\end{equation*}

@@

Random hot deck imputation is a non-parametric method in which missing values in a column are
replaced with randomly selected complete values from the same column, preserving the
distribution of the complete values. We use the same $P$ as for mean value imputation, but
we modify $T$ to account for the fact that hot-deck is blocking in our implementation.  
\begin{equation*}
\begin{aligned}
    P(q) &= \sum_{c \in \cameraready{C}} \frac{1+\widehat{Var}(c)}{\textsc{Card}(q)} \\
    T(q) &= \textsc{Card}(q)
\end{aligned}
\end{equation*}

@@

\begin{figure}
  \cameraready{
  \begin{align*}
    \llbracket Q[T] \rrbracket &= \begin{cases}
      P & (T,\ P) \in Q \\
      \emptyset & \text{otherwise} \\
    \end{cases} \\
    \llbracket Q[T] \lhd p \rrbracket &= \begin{aligned} Q \cup \{(T,\ P) \} \end{aligned} \text{ where }\\
    P &= \left\{q \in S ~|~ \begin{aligned}
              \textsc{Dirty}(q) \neq \textsc{Dirty}(p) \lor
              \not \exists q' \in S:\ \doms{q'}{q}
              \end{aligned}
            \right\} \\
    S &= \llbracket Q[T] \rrbracket \cup \{p\}
  \end{align*}
  \caption{\cameraready{Semantics of the plan cache access and insertion syntax. We use $\llbracket\ \rrbracket$ to mean evaluation. $Q$ is a plan cache, $T$ is a set of tables, and $p$ is a query plan. The cache guaranteees that plans for the same set of tables either have distinct dirty sets or are Pareto optimal.}}
  \label{fig:semantics-plan-cache}
  }
\end{figure}

>>>>

\begin{table}
  \centering
  \caption{\cameraready{Notation used in Section~\ref{sec:algorithm}.}}
\cameraready{
  \begin{tabular}{ll}
    \toprule
    Operator & Description \\
    \midrule
    $\textsc{Attr}(\cdot)$ & Attributes used in a predicate. \\
    $\textsc{Time}(\cdot)$ & Estimated time to run a plan. \\
    $\textsc{Penalty}(\cdot)$ & Imputation quality penalty of a plan. \\
    $\textsc{Card}(\cdot)$ & Estimated cardinality of a plan. \\
    $\widehat{Var}(\cdot)$ & Estimated variance of a column. \\
    $\textsc{Dirty}(\cdot)/\textsc{Clean}(\cdot)$ & Dirty/clean attributes in a plan. \\
    $Q[\cdot]$ & Plan cache access operator. \\
    $\cdot \triangleleft \cdot$ & Plan cache insertion operator. \\
    \bottomrule
  \end{tabular}
}
\end{table}

@@

Separately, we specify a heuristic penalty that takes into account the
improvement in imputation quality from more tuples and attributes, as well as
the diminishing returns from more data.
\[
    P(q) = \frac{|\cameraready{C}|}{\sqrt{ |\textsc{Attr}(q)| \times \textsc{Card}(q)}}
\]

@@

We can take advantage of the histograms maintained by \ProjectName{} to estimate the mean and construct a non-blocking impute operator.
Applying the heuristics from~\Cref{sec:quality} and~\Cref{sec:query-runtime}, we
have
\begin{equation*}
\begin{aligned}
    P(q) &= \sum\limits_{c \in \cameraready{C}} \frac{1+\widehat{Var}(c)}{\textsc{Card}(q)} \\
    T(q) &\approx 0.
\end{aligned}
\end{equation*}

@@

Random hot deck imputation is a non-parametric method in which missing values in a column are
replaced with randomly selected complete values from the same column, preserving the
distribution of the complete values. We use the same $P$ as for mean value imputation, but
we modify $T$ to account for the fact that hot-deck is blocking in our implementation.  
\begin{equation*}
\begin{aligned}
    P(q) &= \sum\limits_{c \in \cameraready{C}} \frac{1+\widehat{Var}(c)}{\textsc{Card}(q)} \\
    T(q) &= \textsc{Card}(q)
\end{aligned}
\end{equation*}

@@

\begin{figure}
  \cameraready{
  \begin{align*}
    \llbracket Q[T] \rrbracket &= \begin{cases}
      P & (T,\ P) \in Q \\
      \emptyset & \text{otherwise} \\
    \end{cases} \\
    \llbracket Q[T] \lhd p \rrbracket &= \begin{aligned} Q \cup \{(T,\ P) \} \end{aligned} \text{ where }\\
    P &= \left\{q \in S ~|~ \begin{aligned}
              \textsc{Dirty}(q) \neq \textsc{Dirty}(p) \lor
              \not \exists q' \in S:\ \doms{q'}{q}
              \end{aligned}
            \right\} \\
    S &= \llbracket Q[T] \rrbracket \cup \{p\}
  \end{align*}
  \caption{\cameraready{Semantics of the plan cache access and insertion syntax. We use $\llbracket\ \rrbracket$ to mean evaluation. $Q$ is a plan cache, $T$ is a set of tables, and $p$ is a query plan. The cache guarantees that plans for the same set of tables either have distinct dirty sets or are Pareto optimal.}}
  \label{fig:semantics-plan-cache}
  }
\end{figure}

### 2017-07-14 paper/conclusion.tex
paper: fix overfull hbox and length problems

\pagebreak
\subsection{Future Work}
\ProjectName{} opens up multiple avenues for further work.

\begin{itemize}
\item \textbf{Imputation confidence:} Obtaining confidence measures for query results
produced when missing values are imputed with a statistical model is
another possible improvement to the system. These measures could be obtained in general
using resampling methods like cross-validation and bootstrapping\cite{kohavi1995study}.
In contrast to standard confidence calculations, one added complexity is that
imputation takes place at various points in the plan, so the composition of results
through standard query operations needs to be taken into account when computing confidence
measures.
In addition, multiple imputation could be used within query execution and confidence measures like variances could
similarly propagate through the query plan.

@@

\item \textbf{Adaptive query planning:} Currently, the imputation is local to a given query, meaning
no information is shared across queries. An intriguing direction would be to take advantage
of imputation and execution data generated by repeated queries. Note that this goes beyond
simply caching imputation results, and instead could entail operations such as extending
intermediate results with prior imputation values to improve accuracy, or pruning out
query plans which have been shown to produce low-confidence imputations.
\end{itemize}

>>>>

\subsection{Future Work}
\ProjectName{} opens up multiple avenues for further work.

\begin{itemize}
\item \textbf{Imputation confidence:} Obtaining confidence measures for query results
produced when missing values are imputed is
a possible improvement to the system. These measures could be obtained in general
using resampling methods like cross-validation and bootstrapping~\cite{kohavi1995study}.
In contrast to standard confidence calculations,
imputation takes place at various points in the plan, so the composition of results
through standard query operations needs to be taken into account when computing confidence
measures.
In addition, multiple imputation could be used within query execution and confidence measures like variances could
similarly propagate through the query plan.

@@

\item \textbf{Adaptive query planning:} Currently, the imputation is local to a given query, meaning
no information is shared across queries. An intriguing direction would be to take advantage
of imputation and execution data generated by repeated queries. This goes beyond
simply caching imputations, and instead could entail operations such as extending
intermediate results with prior imputation values to improve accuracy, or pruning out
query plans which have been shown to produce low-confidence imputations.
\end{itemize}

### 2017-07-14 paper/relatedwork.tex

\section{Related Work}

@@

\subsection{Missing Values and Databases}
There is a long history in the database community surrounding the
treatment of \nullv{}.
Multiple papers have described various (and at times conflicting) treatments of \nullv{}s~\cite{codd1973understanding,grant1977null}.
\ProjectName's design invariants eliminate
the need to handle \nullv{} value semantics, while guaranteeing query evaluation soundness.

@@

\textcite*{duan2007processing} describe the Fa system and its declarative language for time-series forecasting. Their
system automatically searches the space of attribute combinations/transformations and statistical models
to produce forecasts within a given accuracy threshold. Accuracy estimates are determined using
standard techniques, such as cross-validation. 
As with Fa, \ProjectName{} provides a declarative language, as
a subset of standard SQL.\@ \ProjectName{}, however, is not searching the space of possible
imputation models, but rather the space of query plans that incorporate imputation operators. Another major point
of distinction between \ProjectName{} and Fa is that the latter doesn't consider trade-offs between accuracy and computation time, but rather returns the most accurate forecast (with some stopping criterion).

>>>>

\section{Related Work}\label{sec:relatedwork}

@@

\subsection{Missing Values and Databases}\label{subsec:missing-value}
There is a long history in the database community surrounding the
treatment of \nullv{}.
Multiple papers have described various (and at times conflicting) treatments of \nullv{}s~\cite{codd1973understanding,grant1977null}.
\ProjectName's design invariants eliminate
the need to handle \nullv{} value semantics, while guaranteeing query evaluation soundness.

@@

\textcite*{duan2007processing} describe the Fa system and its declarative language for time-series forecasting. Their
system automatically searches the space of attribute combinations/transformations and statistical models
to produce forecasts within a given accuracy threshold. Accuracy estimates are determined using
standard techniques, such as cross-validation. 
As with Fa, \ProjectName{} provides a declarative language, as
a subset of standard SQL.\@ \ProjectName{}, however, is not searching the space of possible
imputation models, but rather the space of query plans that incorporate imputation operators,
and considers the trade-offs between accuracy and computation time.
